---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SCEG-HiC (Predicting enhancer-gene links from single-cell multi-omics data by integrating prior Hi-C information)

<!-- badges: start -->
<!-- badges: end -->
## Overview
SCEG-HiC predicts enhancer–gene links by integrating multi-omics single-cell data (either paired scATAC-seq/RNA-seq or scATAC-seq alone) with three-dimensional chromatin conformation information derived from bulk average Hi-C data. It employs the weighted graphical lasso (wglasso) model to incorporate average bulk Hi-C data, effectively regularizing the correlation matrix with the prior Hi-C contact matrix as a penalty term.

## Installation

### Required software

SCEG-HiC runs in the [R statistical computing environment](https://www.r-project.org/). It requires R version 4.1.0 or higher, [Bioconductor](https://bioconductor.org/) version 3.14 or higher, and Seurat 4.0 or higher to access the latest features.

To install Bioconductor, open an R session and run:

``` r
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.14")
```

Next, install a few Bioconductor packages that are not installed automatically:

``` r
BiocManager::install(c(
  'BiocGenerics', 'DelayedArray', 'DelayedMatrixStats',
  'limma', 'lme4', 'S4Vectors', 'SingleCellExperiment',
  'SummarizedExperiment', 'batchelor', 'HDF5Array',
  'terra', 'ggrastr', 'Gviz', 'rtracklayer', 'GenomeInfoDb', 'GenomicRanges'
))
```

Installation of other dependencies

- Install the Signac pacakge: ```devtools::install_github("timoast/signac", ref = "develop")```. If you encounter any issues, please check the [Signac documentation](https://stuartlab.org/signac/).
- Install the Cicero package: ```devtools::install_github("cole-trapnell-lab/cicero-release", ref = "monocle3")```. If you encounter any issues, please check the  [Cicero installation guide](https://cole-trapnell-lab.github.io/cicero-release/docs_m3/#installing-cicero).

Now, you can install the development version of SCEG-HiC from [GitHub](https://github.com/) with:

``` r
# If you haven't installed devtools yet, uncomment and run:
# install.packages("devtools")

# Install the development version of SCEG-HiC from GitHub
devtools::install_github("wuwei77lx/SCEGHiC")
```

If you prefer the stable release version from CRAN, run:

```r
# Install the released version from CRAN
install.packages("SCEGHiC")
```
### Testing the installation

To verify that SCEG-HiC installed correctly, start a new R session and run:

```{r message=FALSE,warning=FALSE}
library(SCEGHiC)
```

If no errors appear, the package is successfully loaded and ready to use.


## Quickstart

This basic example demonstrates how to analyze paired scATAC-seq/RNA-seq data using SCEG-HiC:

```{r,dpi=150, dev='ragg_png', fig.align='center'}
library(SCEGHiC)
library(Signac)

# Load example multi-omics dataset
data(multiomic_small)

# Preprocess the data (aggregation)
SCEGdata <- process_data(multiomic_small, k_neigh = 5, max_overlap = 0.5)

# Define genes of interest
gene <- c("TRABD2A", "GNLY", "MFSD6", "CTLA4", "LCLAT1", "NCK2", "GALM", "TMSB10", "ID2", "CXCR4")

# Get path to example average Hi-C data
fpath <- system.file("extdata", package = "SCEGHiC")

# Calculate Hi-C based weights for enhancer-gene pairs
weight <- calculateHiCWeights(SCEGdata, species = "Homo sapiens", genome = "hg38", focus_gene = gene, averHicPath = fpath)

# Run the SCEG-HiC model
results_SCEGHiC <- Run_SCEG_HiC(SCEGdata, weight, focus_gene = gene)

# Arc plot visualization predicted enhancer-gene links for CTLA4
connections_Plot(results_SCEGHiC, species = "Homo sapiens", genome = "hg38", focus_gene = "CTLA4", cutoff = 0.01, gene_anno = NULL)
# Load fragment data for coverage plotting
frag_path <- system.file("extdata", "multiomic_small_atac_fragments.tsv.gz", package = "SCEGHiC")
frags <- CreateFragmentObject(path = frag_path, cells = colnames(multiomic_small))
Fragments(multiomic_small) <- frags

# Coverage plot and visualize the links of CTLA4
coverPlot(multiomic_small, focus_gene = "CTLA4", species = "Homo sapiens", genome = "hg38",
          assay = "peaks", SCEG_HiC_Result = results_SCEGHiC, SCEG_HiC_cutoff = 0.01)
```

<details>
  <summary>**Session Info**</summary>

```{r}
sessionInfo()
```

</details>


See the [documentation website](https://wuwei77lx.github.io/SCEGHiC/) for more information!


## The bulk average Hi-C data
The human cell types used for averaging are: GM12878, NHEK, HMEC, RPE1, THP1, IMR90, HUVEC, HCT116, K562, and KBM7.

The mouse cell types used for averaging are: two embryonic stem cell types (mESC1, mESC2), CH12LX, CH12F3, fiber, epithelium, and B cells.

The bulk average Hi-C data can be generated using the [Activity by Contact (ABC)](https://www.nature.com/articles/s41588-019-0538-0) model's ```makeAverageHiC.py``` script.

### Download Links for Average Hi-C Data

* **Human bulk average Hi-C (~54.1 GB)**

  Download from: [ENCFF134PUN.bed.gz](https://www.encodeproject.org/files/ENCFF134PUN/@@download/ENCFF134PUN.bed.gz).

After downloading, extract the human bulk average Hi-C using the [Activity by Contact (ABC)](https://www.nature.com/articles/s41588-019-0538-0) model's ```extract_avg_hic.py``` script:

```{sh eval=FALSE}
python code/Hi_C/extract_avg_hic.py --avg_hic_bed_file ../ENCFF134PUN.bed.gz --output_dir ../
```

* **Mouse bulk average Hi-C (~13.9 GB)**

  Download from: [10.5281/zenodo.14849886](https://zenodo.org/record/14849886).


For more details about bulk average Hi-C data, please visit: https://github.com/wuwei77lx/compare_model.

## Example
In SCEG-HiC, you can choose either aggregation or single-cell retention approach:

- **Aggregation approach**: Aggregates binarized scATAC-seq data across cell types using k-nearest neighbor smoothing to reduce data sparsity. This approach captures a broader spectrum of enhancer-gene links across cell types, with slightly reduced prediction accuracy.

- **Single-cell retention approach**: Normalizes scATAC-seq data within individual cell types to individual cell signals. This method achieves higher precision and accuracy, albeit identifying fewer enhancer-gene links

**Recommendation**: To balance accuracy and coverage, we implemented both preprocessing strategies in SCEG-HiC, with aggregation designated as the default. The single-cell retention approach can be optionally used when higher precision within specific cell types is desired.

For more details and real data examples, please visit:

- [SCEG-HiC on paired scATAC-seq/RNA-seq data of PBMC (aggregation)](https://wuwei77lx.github.io/SCEGHiC/articles/PBMC_multiomic_aggregation.html)

- [SCEG-HiC on paired scATAC-seq/RNA-seq data of PBMC CD4T](https://wuwei77lx.github.io/SCEGHiC/articles/PBMC_multiomic_CD4T.html)

- [SCEG-HiC on paired scATAC-seq/RNA-seq data of mouse skin (aggregation)](https://wuwei77lx.github.io/SCEGHiC/articles/mouse_skin_multiomic_aggregation.html)

- [SCEG-HiC on scATAC-seq data alone from human COVID-19 monocytes](https://wuwei77lx.github.io/SCEGHiC/articles/human_covid_19_scATAC_seq_monocytes.html)

## Help

If you have any questions, comments, or suggestions, please contact Xuan Liang at liangxuan2022@sinh.ac.cn.
