---
title: "SCEG-HiC on paired scRNA-seq and scATAC-seq data of PBMC-CD4T"
author: "XuanLiang"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/picb/bigdata/project/liangxuan/data/human/pbmc")
options(future.globals.maxSize = 20 * 1024^3)
```

In this vignette,we’ll demonstrate  SCEG-HiC’s capability of inferring gene-enhancer by applying it to paired scRNA-seq and scATAC-seq data. In this vignette we'll be using a publicly available 10x Genomic Multiome dataset for human PBMCs,and only select CD4 T cells for processing

Ok, so let’s get started. First we need a Seurat object that has paired scRNA-seq and scATAC-seq data, respectively. Here we use multiome data from human PBMCs, processed as demonstrated in the [SCEG-HiC on paired scRNA-seq and scATAC-seq data of PBMC-aggregate](https://wuwei77lx.github.io/SCEGHiC/articles/pbmc_multiomic_aggregrate.html).

The implementent of SCEG-HiC is seamlessly compatible with the workflow of Seurat/Signac package, SCEG-HiC pipeline consists of the following four major parts:

* Infer gene-enhancer using SCEG-HiC by taking as input input a Seurat object and the bulk average Hi-C.
* Visualize the gene-enhancer.

## Load the required libraries
```{r message=FALSE, warning=FALSE}
library(SCEGHiC)
library(Signac)
library(Seurat)
library(EnsDb.Hsapiens.v86)
library(BSgenome.Hsapiens.UCSC.hg38)
library(ggplot2)
```

## Infer gene-enhancer using SCEG-HiC 

### Preprocess data
Here, we only select CD8 T cells with paired scRNA-seq and scATAC-seq data.By default, the scATAC-seq counts are binarized.
```{r}
# load data
load("pbmc_multiomic.rda")
SCEGdata <- process_data(pbmc, aggregate = FALSE, celltype = "CD4 T", rna_assay = "SCT")
```
### Calculate weight
SCEG-HiC takes a long time to calculate weight, and the process becomes slower as more genes are selected. SCEG-HiC provides download websites for the average Hi-C data for both humans and mice. In this case, we select the human average Hi-C data for analysis.

```{r }
# For example, Calculate weight for CD4 T's markers: GLB1
weight <- calculateHiCWeights(SCEGdata, species = "Homo sapiens", genome = "hg38", focus_gene = "GLB1", averHicPath = "/picb/bigdata/project/liangxuan/data/human_contact/AvgHiC")
```
### Run model
SCEG HiC uses wglasso machine learning method and normalizes the bulk average Hi-C matrix using rank score.
```{r }
# For example, Run model for CD4 T's markers: GLB1
results_SCEGHiC <- Run_SCEG_HiC(SCEGdata, weight, focus_gene = "GLB1")
```

## Visualize the gene-enhancer 

### Visualize the links

```{r,fig.width = 8, fig.height = 5,message=FALSE,warning=FALSE}
temp <- tempfile()
download.file("https://hgdownload2.soe.ucsc.edu/goldenPath/hg38/bigZips/genes/hg38.refGene.gtf.gz", temp)
gene_anno <- rtracklayer::readGFF(temp)
unlink(temp)
# rename some columns to match requirements
gene_anno$chromosome <- gene_anno$seqid
gene_anno$gene <- gene_anno$gene_id
gene_anno$transcript <- gene_anno$transcript_id
gene_anno$symbol <- gene_anno$gene_name
connections_Plot(results_SCEGHiC, species = "Homo sapiens", genome = "hg38", focus_gene = "GLB1", cutoff = 0.001, gene_anno = gene_anno)
```

<details>
  <summary>**Session Info**</summary>

```{r}
sessionInfo()
```

</details>
